Multi-modality Imagery Database for Plant Phenotyping

### 2015-05-04 datapaper_calibration.tex
noise modeling clarifications

The depth noise $\epsilon$ is modeled as the sum of two components, $\epsilon_I$ and $\epsilon_S$:
\begin{equation}
\epsilon = \epsilon_I + \epsilon_S\label{eq:epsilon}
\end{equation}
The term $\epsilon_I$ captures the noise component that varies between images taken from a fixed pose of a static scene. On the other hand the term $\epsilon_S$ models noise is constant between images but varies between pixels. The variance of $\epsilon_I$ is estimated for each pixel of a fixed scene observed over multiple images. In our experiments we observed a flat, uniform albedo surface perpendicular to the camera at a sequence of depths, and for each depth acquired 300 images. Object depth, $d$, has a large impact on $\epsilon_I$ likely due to reduced reflected signal. For constant depth, we observed that the primary factor affecting the variance is the pixel radius, $r$, from the image center. Physically we expect this dependence is due to a circularly symmetric illuminating beam closely aligned with the optical axis. Based on these observations we model $\sigma_I^2(d,r)$ the variance of $\epsilon_I$, as a function of depth and pixel radius.  From our experiments we build up a lookup table for this as plotted in Figure~\ref{fig:Noise}.

@@

In addition to repeated-image-varying noise, there are pixel depth offsets that are constant for images of the same scene, and so cannot be removed by averaging over many frames. We model these with $\epsilon_S$. To estimate its variance, $\sigma_S^2$, we first average over many depth measurement images, in our case 300, to obtain pixel depth estimates that approximately eliminate the effect of $\epsilon_I$. Then by calculating true ray depths on a known surface, in our case the observed plane, and assuming that $\epsilon_S$ has the same distribution over all pixels, $\sigma_S$ is obtained as the standard deviation of the error between averaged depths and known depths. In our experiments we obtained $\sigma_S=6.5mm$, and found that it was insensitive to changes in depth.

The recorded depth images in the data collection are the result of averaging $N=5$ subsequent depth images.  Assuming independence of $\epsilon_I$ and $\epsilon_S$, the expected variance of pixel depth measurements for is given by:
\begin{equation}
\sigma^2(d,r) = \frac{\sigma_I^2(d,r)}{N} + \sigma_S^2.\label{eq:sigma}
\end{equation}

There are additional sources of noise not captured by this including object specularities, sharp variations in object albedo, mixed-depth pixels on object edges, and cases of very-low signal reflection. One of the utilities of having a modeled expected variance is that it can be compared with the measured variance, and strong disparity is a cue that the surface violates our modeling assumptions, and can be filtered out as not belonging to a plant leaf.

In addition we noticed that the chamber light shades blocked some of the depth camera field of view, and in doing so reflected some of the IR illumination. This resulted in a small constant depth shift for the pixels. We measured this shift for each chamber experiment and provide it as an optional correction to the depth images.

>>>>

The depth noise, $\varepsilon$, is modeled as the sum of an image dependent term, $\varepsilon_I$, and a sensor dependent term, $\varepsilon_S$:
\begin{equation}
\varepsilon = \varepsilon_I + \varepsilon_S\label{eq:epsilon}
\end{equation}
The term $\varepsilon_I$ is a random variable for each pixel with a value that varies between subsequent images taken from a fixed pose of a static scene.  On the other hand, $\varepsilon_S$ is a random variable for each pixel that models its depth offset, and its value only changes when the scene changes.  The variance of $\varepsilon_I$ is estimated for each pixel of a fixed scene observed over multiple images. In our experiments we observed a flat, uniform albedo surface perpendicular to the camera at a sequence of depths, and for each depth acquired 300 images. Object depth, $d$, has a large impact on $\varepsilon_I$. For constant depth, we observed that the primary factor affecting the variance is the pixel radius, $r$, from the image center. Physically we expect this dependence is due to a circularly symmetric illuminating beam closely aligned with the optical axis. Based on these observations we model $\sigma_I(d,r)$, the standard deviation of $\varepsilon_I$, as a function of depth and pixel radius.  From our experiments we build up a lookup table for this as plotted in Figure~\ref{fig:Noise}.

@@

Averaging over repeated images of a scene will not remove all the depth error as there are pixel depth offsets that are constant for images of the same scene. We model these with $\varepsilon_S$. To estimate its standard deviation, $\sigma_S$, we first average over many depth measurement images, in our case 300, to obtain pixel depth estimates that approximately eliminate the effect of $\varepsilon_I$. Then by calculating true ray depths on a known surface, in our case the observed plane, and further assuming that $\varepsilon_S$ has the same standard deviation for all pixels, $\sigma_S$ is obtained as the standard deviation of the error between averaged depths and known depths. In our experiments we obtained $\sigma_S=6.5mm$, and found that it was insensitive to changes in depth.

The recorded depth images in the data collection are the result of averaging $N=5$ subsequent depth images.  Assuming independence of $\varepsilon_I$ and $\varepsilon_S$, the variance of pixel depth measurements is given by:
\begin{equation}
\sigma^2(d,r) = \frac{\sigma_I^2(d,r)}{N} + \sigma_S^2.\label{eq:sigma}
\end{equation}

There are additional sources of noise not modeled by this.  Object albedo has an impact although this is fairly weak for strong signal reflections.  Factors with large impact on signal noise include: object specularities, sharp variations in object albedo, mixed-depth pixels on object edges, and cases of very-low signal reflection, all of which can lead to very large variances. One of the utilities of having a model for variance is that it can be compared with the measured variance, and the difference used as a cue for portions of the scene that violate our modeling assumptions.

In addition, we noticed that the chamber light shades blocked some of the depth camera field of view, and in doing so reflected some of the IR illumination. This resulted in a small constant depth shift for the pixels. We measured this shift for each chamber experiment and provide it as an optional correction to the depth images.

### 2015-05-04 datapaper_hardware.tex
update_Xi
noise modeling clarifications

\begin{figure}
\centering
  \includegraphics[height=5.9cm,trim=110 250 60 260,clip]{Figures/SigmaRadius} \\
\caption{Noise analysis for a depth camera obtained by imaging a flat surface at various depths.  We found that the standard deviation, $\sigma_I$ from Eq. (\ref{eq:sigma}) of the pixel depth measurements had a large dependency on both the target depth and the pixel distance from the image center, and these are plotted.  A radius of $0$ pixels is the image center, and of $200$ pixels corresponds to the corners of the depth camera, which can be seen to have far larger standard deviation than at the image center for the same depth. }
\label{fig:Noise}
\end{figure}

>>>>

\begin{figure}
\centering
  \includegraphics[height=5.9cm,trim=110 250 60 260,clip]{Figures/SigmaRadius} \\
\caption{Noise analysis for a depth camera obtained by imaging a flat surface at various depths.  We found that the standard deviation, $\sigma_I(d,r)$ from Eq. (\ref{eq:sigma}) of the pixel depth measurements had a large dependency on both the target depth, $d$, and the pixel radius, $r$, from the image center, and these are plotted.  A radius of $0$ pixels is the image center, and of $200$ pixels corresponds to the corners of the depth camera, which can be seen to have far larger standard deviation than at the image center for the same depth. }
\label{fig:Noise}
\end{figure}
